---
title: "biomaRt_create_flatfile.Rmd"
date: "July 28, 2020"
output: html_document
---

https://github.com/grimbough/biomaRt/issues/20

due to a confluence of upgrades, R 4.0.* , Ubuntu 18.04, and switching to Rstudio server commercial product, I can no longer query biomaRt regularly. 
could mayybe due biomaRt have increased demand? but basically useEnsembl() still fail regulary with curl errors, and the query also times out with curl errors. 

At this point, my success rate is like 1/4 for actually create a mart object and then completing a query for a gene list. 

#Set-up

```{r setup, cache = FALSE, include = FALSE}
require(knitr)
knitr::opts_knit$set(root.dir = file.path(SCRIPTS))
```

```{r eval=FALSE}
library(biomaRt) #lib.loc = lib.path


#Having a hella errors when running these in the function itself. 
mart.37 <- useEnsembl(biomart = "ENSEMBL_MART_ENSEMBL",
                        dataset = "hsapiens_gene_ensembl",
                        GRCh = 37)   

# saveRDS(mart.37, file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/biomaRt.GRCh37.RDS"))
```

```{r eval=FALSE}
mart.38 <- useEnsembl("ensembl",
                       mirror = "www",
                       dataset = "hsapiens_gene_ensembl")


# saveRDS(mart.38,file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/biomaRt.GRCh38.RDS"))
```

```{r eval=FALSE}
all_gene_ids.37 <- biomaRt::getBM(attributes = c("ensembl_gene_id"),
                               mart = mart.37) #63677
all_gene_ids.38 <- biomaRt::getBM(attributes = c("ensembl_gene_id"),
                               mart = mart.38) #67149

# saveRDS(all_gene_ids.37,file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/GRCh37_all_gene_IDs.RDS"))
# saveRDS(all_gene_ids.38,file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/GRCh38_all_gene_IDs.RDS"))
```

```{r}
mart.37 <- readRDS(file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/biomaRt.GRCh37.RDS"))

mart.38 <- readRDS(file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/biomaRt.GRCh38.RDS"))

all_gene_ids.37 <- readRDS(file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/GRCh37_all_gene_IDs.RDS"))

all_gene_ids.38 <- readRDS(file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/GRCh38_all_gene_IDs.RDS"))
```

#Attributes 

```{r}
attr.mart <-  c("ensembl_gene_id",
                  "external_gene_name",
                  "transcript_count",
                  "ensembl_transcript_id", 
                  "ensembl_peptide_id",
                  "tmhmm", "tmhmm_start", "tmhmm_end") #"transcript_tsl",
```


```{r}

```

#GRCh37 results

```{r}
res.anno.37 <- getBM(attributes = attr.mart,
                    filters = "ensembl_gene_id",
                    values = all_gene_ids.37$ensembl_gene_id[1:50],
                    mart = mart.37)

res.anno.37
```
  

#GRCh38 results  
```{r}  

res.anno.38 <- getBM(attributes = attr.mart,
                     filters = "ensembl_gene_id",
                     values = all_gene_ids.38,
                     mart = mart.38)


# saveRDS(res.anno.38, file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/biomaRt.GRCh38_allGenes.RDS"))
# saveRDS(res.anno.37, file.path(PROJHOME,"0000.00.02_Reference_GeneInfo/biomaRt.GRCh37_allGenes.RDS"))
```
