---
title: "Concatenate 1031 Exon Coverage"
output: html_document
---

#Set-up

```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center', fig.height = 8, fig.width = 10)
knitr::opts_knit$set(root.dir = '/fh/fast/meshinchi_s/workingDir/TARGET/AML_TARGET/RNA/mRNAseq/analysis/2017.10.09_Concatenate_1031_RNAseq/')
```

```{r message=FALSE}
library(dplyr)
library(magrittr)
library(ggplot2)
library(stringr)
library(reshape2)
library(readr)
getwd()
```

```{r}
source("~/scripts/conversion_scripts/Merge_Cat_FixDupIDs_Function.r")
```

```{r}
addCols <- function(df,symbol,exon){
  library(dplyr)
  
  df <- as.data.frame(df)
  
  df <- df %>%
    mutate(geneSymbol=symbol,
           exon=exon) %>%
    select(geneSymbol, exon, everything())
  
  return(df)
}
```


```{r}
#Function for the TPM conversion. 
# Based on https://groups.google.com/forum/#!topic/rsem-users/W9RQrZIOzA4
RPKM_to_TPM <- function(RPKM){
  conversionFactor <- sum(RPKM) / 1E6
  TPM <- RPKM / conversionFactor
  return(TPM)
}
```


#Identify Files to Be Concatenated

```{r}
filepath <-  "/fh/fast/meshinchi_s/workingDir/TARGET/AML_TARGET/RNA/mRNAseq/level3/exon/2017July_BCCA_0531_1031_Ribodepletion_Illumina_data/"

allfiles <- dir(path = filepath,
                pattern = ".normalized")

head(allfiles)
```


```{r}
length(allfiles) #1,116 files (no stella)
```

```{r}
#target matrix will have NBM, AML, and untreated cell line samples ("D1" == day 1)
target <- paste0(filepath, grep("^[RBPS][A-Z0-9\\-]", allfiles, value=TRUE)) #1,111 samples

#cell line matrix will have Kasumis and MV411s treated and untreated.  
cells <- paste0(filepath, grep("^Kas|MV4", allfiles, value=TRUE))  #6 samples
```

```{r}
head(target, n=2)
head(cells, n=2)
```




#Columns Description

The .exon.normalized file contains: For BCCA
-*column 1*: gene id
-column 2: transcript id or "merged_<gene_id>" if the analysis was performed in
collapse mode
-colomn 3: strand of the annotation ("+" or "-")
-*column 4*: exon number (relative to the gene's strand)
-column 5: "exon" (annotation type)
-column 6: chromosome
-column 7: start of the exon (relative to the positive strand, so start<end)
-column 8: end of the exon (relative to the positive strand, so start<end)
-*column 9*: number of fractional reads inside this exon's boundaries (sum of the
fraction of each read inside this region)
-column 10: total coverage across this exon (sum of the coverage depth at each
base in this exon)
-column 11: average coverage inside this exon-- total coverage across this exon
divided by the exon's length
-*column 12*: normalized average coverage inside this exon (RPKM) -- (number of this
exon's fractional reads x 1000000000) / (NORM_TOTAL x exon length)


#Begin Concatenation 


```{r}
#Pattern to select the Target Barcode
pattern <- "^.+\\/([BPR][A-Z0-9].+R)\\_.+"
pattern.cells <- ".+\\/([KM][A-Za-z0-9].+R)\\_.+"
```

```{r}
#Select the column indices 
selected <- c(1,4,9,12)
```

```{r results='hide'}
filesWithPath <- list(target[1:10], cells)
patterns <- list(pattern, pattern.cells)

cated <- mapply(catExpnData,filesWithPath, patterns, MoreArgs = list(cols=selected,header=FALSE), SIMPLIFY = FALSE)

names(cated) <- c("TARGET", "Cells")
```

```{r}
sapply(cated, function(x) sapply(x,dim)) %>% set_rownames(rep(c("genes", "samples"), 4))
```

```{r}
gc()
# save(cated, file="TARGET_AML_1031_ExonLevel_RNAseq_Cated.RData")
load("TARGET_AML_1031_ExonLevel_RNAseq_Cated.RData")
```


#Convert RPKM to TPM 

```{r}
options(scipen = 999)

for (i in 1:2){
  names(cated[[i]]) <- c("geneSymbol", "Exon", "FractionalCounts", "RPKM")
  head(cated[[i]]$RPKM)
  cated[[i]][["TPM"]] <- apply(cated[[i]]$RPKM,2, RPKM_to_TPM)
}

```

```{r}
IDmap <- read.csv("GeneSymbol_EnsemblID_Conversion.csv")
head(IDmap)
```


#Check that the Gene Ids are in the Same Order

```{r}
# apply(cated$TARGET$`1`, MARGIN=2,FUN=identical, y=cated$TARGET$`1`[,1])
sapply(cated$TARGET, function(mat) all(apply(mat, MARGIN=2,FUN=identical, y=mat[,1])))

```
Note: Columns are the counts, so would not expect or want these to be identical for all samples (columns)

```{r}
sapply(cated$Cells, function(mat) all(apply(mat, MARGIN=2,FUN=identical, y=mat[,1])))
```



#Add columns for the Gene Symbol and Ensembl Symbol 

"FractionalCounts", "RPKM"
```{r}
withCols <- lapply(cated$TARGET[3:5], addCols, 
                   symbol = cated$TARGET$geneSymbol[,1], exon = cated$TARGET$Exon[,1])


lapply(withCols, function(x) head(x[,1:5]))
```

```{r}
withCols.cells <- lapply(cated$Cells[3:5], 
                         addCols, symbol = cated$Cells$geneSymbol[,1], exon = cated$TARGET$Exon[,1])

lapply(withCols.cells, function(x) head(x[,1:5]))
```


#Save the output

```{r}
list <- list(withCols, withCols.cells)
names(list) <- c("TARGET", "Cells")

for (i in 1:2){
  n <- ifelse(names(list[i]) == "TARGET", "TARGET_AML_AAML1031_RNASeq_Exon_", "Kasumi_MV4-11_Exon_")
  
  for (idx in names(list[[i]])){
    filename <- paste0(n, idx,".csv")
    print(filename)
    write.csv(list[[i]][[idx]], filename, row.names = FALSE)
  }
}

```



#Create Subset of CBL expression 

```{r}
TPMs  <- read_csv("exonLevel/BCCA_inHouse_Pipeline/TARGET_AML_AAML1031_RNASeq_Exon_TPM.csv")

dim(TPMs) #332,261   1113
head(TPMs[,1:5])
```

CBL:	ENSG00000110395

```{r}
CBL <- TPMs %>% 
  select(gene_id=geneSymbol, everything()) %>%
  filter(gene_id=="ENSG00000110395") %>% 
  gather(USI,TPMs, -gene_id, -exon) %>%
  mutate(gene_name="CBL",
         transcript_id="ENST00000264033") %>%
  mutate(Group=case_when(grepl("^BM|^RO", USI) ~ "NBM",
                         grepl("^P", USI) ~ "AML",
                         grepl("Kas|MV4", USI) ~ "Celllines"))


head(CBL)
dim(CBL)
```

```{r}
write.csv(CBL,"exonLevel/BCCA_inHouse_Pipeline/TARGET_AML_AAML1031_CBL_Exon_TPMs.csv", row.names = FALSE)
```

```{r}
table(CBL$Group)
table(CBL$exon)
```







NOTE:

If markdown cannot be completed with pandoc (memory issues), use the following

> library("rmarkdown")
> render("~/scripts/RNAseq_Analysis/DifferentialExpn_PathwayAnalysis/TARGET_AML_1031_Cat_RNAseq.Rmd", "html_document")


#Session Info

```{r}
sessionInfo()
```