---
title: 'Concatenate STAR Fusion for St Jude AML and ALL Samples'
author: "Jenny Smith"
date: "Oct 2, 2020"
output: html_document
---

# Set-up

```{r setup}
library(knitr)
# dir.create(file.path(PROJHOME,"2020.10.09_St.Jude_STAR_Fusion"))
knitr::opts_knit$set(root.dir = file.path(PROJHOME,"2020.10.09_St.Jude_STAR_Fusion"))
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, 
                      fig.align='center', 
                      fig.width = 10, fig.height = 10)

options(stringsAsFactors = FALSE)
```

```{r message = FALSE, warning=FALSE}
library(stringr)
library(magrittr)
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
getwd()
```

```{r}
source(file.path(SCRIPTS,"conversion_scripts/Merge_Cat_FixDupIDs_Function.r"))
```


#Define Functions to be used

```{r}
collapseRows <- function(col, uniq=FALSE){
  #designed for dplyr so that col is the vector of the column name
  if (uniq){col <- unique(col)}
  
  collapsed <- ifelse(all(is.na(col)), NA, paste(col, collapse = "; "))
  return(collapsed)
}


removeDups <- function(col.rowwise){
  
  uniqued <- str_split(col.rowwise, "; ") %>% 
  unlist() %>%
  gsub(" ","",.) %>%
  unique() %>%
  paste(.,collapse = "; ")
  
  return(uniqued)
  
}

```


```{r}
fusionCategory <- function(geneA,geneB){
  fus <- c(geneA,geneB)
  fus <- paste(fus[order(fus)], collapse = "-") #order alphabetically, so that the category includes  reciprocal fusions
}
```

```{r}
collapseFusions <- function(breakpoint){
  paste(unique(breakpoint), collapse = "; ")
}
```

```{r}
fillNA <- function(x,USI){
  
  USI <- unique(USI)
  x <- x %>% .[!is.na(.)] %>% unique()
  
  if(length(x) < 1){
    x <- ""
  }
  
  # print(x)
  return(x)
}
```

```{r}
getPrimaryFusion <- function(USI,junction_reads,spanning_reads){


  resolveTies <- function(df){
        df <- df %>%
            group_by(primary) %>%
            #if still a tie in breakpoint pairs counts, choose the first.
            mutate(rank=rank(junction_reads, ties.method = "first")) %>%  
            #convert all alternate fusion calls to zero rank (eg not going to be used in tie-breaking)
            mutate(rank=ifelse(primary=="alternate", 0, rank)) %>%
            rowwise() %>%
            mutate(reason=sum(reason,rank)) %>%
            ungroup() %>% 
            #the fusion with highest # of breakpoint reads will have largest interger rank. 
            mutate(primary=ifelse(reason == max(reason),
                                  "primary", "alternate"))
        
        return(df)
    }
  
  #First check that the reads are not NA values (eg. NBM samples)
  if (all(is.na(junction_reads))){
    res <- NA
    return(res)
  }
  
  
  #Create boolean vectors for each type of evidence supporting the "primary" fusion. 
  USI <- unique(USI)
  brk.reads <-  junction_reads == max(junction_reads)
  spanning.reads <- spanning_reads == max(spanning_reads)

  #Create a temporary data frame. 
  temp <- data.frame("brk.reads.bool"=brk.reads,
                     "spanning.reads.bool"=spanning.reads, 
                     stringsAsFactors = FALSE) %>%
    #add in the subsetted per Sample read counts. For visual checking on ties, etc. 
    bind_cols(.,list("junction_reads"=junction_reads,
                     "spanning_reads"=spanning_reads))
  
  #will select the one with the highest reads
    temp <- temp %>%
        rowwise() %>%
        mutate(reason=sum(brk.reads.bool,spanning.reads.bool)) %>%
        ungroup() %>%
        mutate(primary=ifelse(reason == max(reason), "primary", "alternate"))

    # break ties if there are multiple fusions with same #s of supporting reads.
    # for those with highest evidence ("reason"), choose the one with the higher breakpoint reads
    if (sum(temp$primary == "primary") > 1){
        temp <- resolveTies(temp)
        
        # last check for ties, will break dplyr by returning a dataframe, not a vector. 
        if (sum(temp$primary == "primary") > 1 ){
          return(temp)
        }
    }

  res <- temp$primary
  
  return(res)
}
```


#Read in the Manifest

```{r}
manifest <- read.csv(file.path(TARGET,"SequencingDataMatrix/St.Jude_AML_ALL_RNAseq_DUA_Manifest.csv"))


head(manifest)
dim(manifest)
```

```{r}
any(duplicated(manifest$sample_name))
table(duplicated(manifest$subject_name))
```



# Download the files from S3

```{bash eval=FALSE}
cd "/fh/scratch/delete90/meshinchi_s/jlsmith3/STAR-Fusion_St.Jude"

aws s3 cp  --recursive --exclude "*" --include "*annotated.coding_effect" --include "*html"  s3://fh-pi-meshinchi-s/SR/starfusion/ .

```


#Concatenate the Sample Level Files 

```{r}
path <- file.path(SCRATCH,"jlsmith3/STAR-Fusion_St.Jude")

files <- dir(path=path,recursive = TRUE,
             full.names = TRUE,
             pattern="*tsv.annotated.coding_effect")

length(files) #487
head(files)
```

```{r}
bam_files <- data.frame(filepath=dir(path=path,recursive = TRUE,
             full.names = TRUE,
             pattern="^.+Aligned.sortedByCoord.out.bam$")) 

head(bam_files)
dim(bam_files)
```

```{r}
bam_manifest <- bam_files %>% 
  mutate(filename=str_split_fixed(filepath, "\\/", n=8)[,8]) %>%
  mutate(sample_name=str_split_fixed(filename, "\\.", n=2)[,1]) %>%
  left_join(., manifest, by="sample_name") %>% 
  select(Sample=sample_name,subject_name:attr_sex, -filepath, -filename, everything(), filepath, filename)

head(bam_manifest)
# write.csv(bam_manifest,
#           file.path(PROJHOME,"2020.09.15_RNAseq_Fusion_Breakpoints/BAM_Manifests/St.Jude_AML_ALL_RNAseq_Bam_Manifest_10.09.20.csv"),
#           row.names = F)
```

```{r}
regex="^.+STAR-Fusion_St.Jude/(SJ.+)\\.RNA-Seq_/Fusion.+"
fusionCaller <- catRbind(filenames=files, regex=regex, header=TRUE, ID.Col.Name="Sample") 

fusionCaller <- fusionCaller %>% 
  mutate(subject_name=str_split_fixed(Sample, "_", n=3)[,1]) %>%
  add_row(Sample=setdiff(bam_manifest$Sample, .$Sample),
          X.FusionName="NoneDetected") %>% 
  select(Sample, subject_name, everything())


dim(fusionCaller) #5786   33
head(fusionCaller[,1:5])
# tail(fusionCaller)
# length(unique(fusionCaller$Sample)) #538
# write.csv(fusionCaller, "St.Jude_AML_ALL_STAR_Fusion_GRCh37.csv", row.names=FALSE)
```



#Read in the STAR Fusion Data

```{r}
star <- read.csv("St.Jude_AML_ALL_STAR_Fusion_GRCh37.csv", 
                 stringsAsFactors = FALSE) %>% 
  mutate(subject_name=str_split_fixed(Sample, "_", n=3)[,1]) 

length(unique(star$Sample)) # 491
dim(star) #5786   33
head(star)
# tail(star)
```



#Reformat STAR fusion to Include Custom Columns 

```{r}
star.fmt <- star %>% 
    arrange(Sample) %>%
    select(Sample, X.Fusion=X.FusionName,
           Left.Breakpoint=LeftBreakpoint,
           Right.Breakpoint=RightBreakpoint,
           JunctionRead=JunctionReadCount,
           SpanningRead=SpanningFragCount,
          everything()) %>%
    #replace dashes with a dot. for consistency withother datasets.
    mutate_at(vars(Sample,X.Fusion,LeftGene,RightGene),
              ~gsub("-", "\\.", .)) %>% 
    mutate(Left.Gene=str_split_fixed(LeftGene,"\\^",n=2)[,1],
           Right.Gene=str_split_fixed(RightGene,"\\^",n=2)[,1],
           Left.Chr=str_split_fixed(Left.Breakpoint, ":", n=3)[,1],
           Right.Chr=str_split_fixed(Right.Breakpoint,":", n=3)[,1],
           
           Left.pos=as.numeric(str_split_fixed(Left.Breakpoint, ":", n=3)[,2]),
           Right.pos=as.numeric(str_split_fixed(Right.Breakpoint,":", n=3)[,2]),
           
           Left.Strand=str_split_fixed(Left.Breakpoint, ":", n=3)[,3],
           Right.Strand=str_split_fixed(Right.Breakpoint,":", n=3)[,3],
           
           Breakpoints = paste(gsub("chr|:[\\+-]","",Left.Breakpoint),
                               gsub("chr|:[\\+-]","",Right.Breakpoint),
                               sep="|"),
           Type = ifelse(Left.Chr == Right.Chr,
                         "intrachromosomal","interchromosomal"),
           size = ifelse(Type=="intrachromosomal",
                         abs(Left.pos - Right.pos), NA)) %>%
    mutate(X.Fusion = gsub("\\.{2}","-",X.Fusion)) %>%
    mutate(Fusion.Detected.STAR=ifelse(is.na(X.Fusion) | X.Fusion == "NoneDetected",
                                       "NoneDetected","Detected")) %>%
    mutate_at(vars(X.Fusion:Type),
              ~ifelse(Fusion.Detected.STAR=="NoneDetected","NoneDetected",.)) %>%
    select(Sample,subject_name,X.Fusion, Breakpoints,everything())
  

dim(star.fmt)
length(unique(star.fmt$Sample)) 
```

```{r}
head(star.fmt)
# quantile(star.fmt$size, na.rm=T)
# table(star.fmt$Type)
# write.csv(star.fmt, "St.Jude_AML_ALL_STAR_Fusion_reformatted_10.12.20.csv", row.names = FALSE)
```



#Filter out Fusions found in NBMs 

```{r}
filterFusions <- function(X.Fusion, Breakpoints,df2.Fusion,df2.Breakpoints){

  fus <- paste(X.Fusion,Breakpoints,sep=" ")
  # print(fus)
  fus2 <- paste(df2.Fusion, df2.Breakpoints, sep=" ")
  
  if (fus %in% fus2){
      res <- 1 #these fusions will be filtered out. 
  }else{
    res <- 0 #these one will not. 
  }
  
  return(res)
}
```

```{r}
NBM.1031 <- read.csv(file.path(PROJHOME,"2018.02.07_STAR-Fusion_C.Nguyen","Reformatted/TARGET_AML_1031_STAR_Fusion_reformatted_7.02.2018.csv"), 
                     row.names = 1) %>%
  filter(grepl("BM[0-9]|RO[0-9]", USI)) %>% 
  mutate_all(~as.character(.))


head(NBM.1031)
dim(NBM.1031) #5707   17
```

```{r}
# setdiff(colnames(NBM.1031), colnames(star.fmt)) #size only
setdiff(colnames(star.fmt),colnames(NBM.1031))
```

```{r}
NBM.0531 <- read.csv(file.path(PROJHOME,"2018.02.07_STAR-Fusion_C.Nguyen","Reformatted/TARGET_AML_0531_STAR_Fusion_reformatted_4.09.2019.csv")) %>%
  filter(grepl("BM[0-9]|RO[0-9]", USI)) %>% 
  mutate_all(~as.character(.))


head(NBM.0531)
dim(NBM.0531) #1757   19
```

```{r}
#Using both NBM sets to account for the total fusions found in the NBM dataset
NBM <- read.csv(file.path(PROJHOME,"2020.02.13_STAR-Fusion_Relapse/reformatted/TARGET_AML_0531_1031_Relapse_STAR_Fusion_reformatted_4.22.2020.csv")) %>%
  filter(grepl("BM[0-9]|R[O0][0-9]", USI)) %>%
  mutate_all(~as.character(.)) %>%
  bind_rows(., NBM.1031,NBM.0531) %>% 
  select(-one_of(setdiff(colnames(star.fmt),colnames(NBM.1031))))

# head(NBM)
dim(NBM)  # 7583   45 (119 in CD34+ NBM alone)
length(unique(NBM$Patient)) #100
# write.csv(NBM,"TARGET_Fusions_in_NormalMarrrows_STAR-fusion_4.27.20.csv", row.names = FALSE)
```

```{r}
star.fmt.filter <- star.fmt %>% 
  select(Sample,subject_name,everything()) %>%
  
  rowwise() %>%
  mutate(FailedFilterFusions=filterFusions(X.Fusion, 
                                           Breakpoints,
                                           NBM$X.Fusion,
                                           NBM$Breakpoints)) %>%
  ungroup() %>%
    
  group_by(Sample) %>%
  mutate(AllFail=ifelse(sum(FailedFilterFusions) == n(),"Yes","No")) %>%
  mutate(Fusion.Detected.STAR=ifelse(AllFail == "Yes" ,
                                     "NonePassedFilter",
                                     Fusion.Detected.STAR)) %>%
  ungroup()  %>%
  
  filter(AllFail == "Yes" | FailedFilterFusions == 0) %>%

  rowwise() %>%
  #change all to characters so that numeric cols can be filled below with a charcter string.
  mutate_all(~as.character(.)) %>% 
  mutate_at(vars(X.Fusion:Fusion.Detected.STAR), 
            ~ifelse(Fusion.Detected.STAR =="NonePassedFilter",
                    "NonePassedFilter",  .)) %>%
  ungroup() %>%
  
  select(Sample,subject_name,
         everything(),-AllFail,-FailedFilterFusions) %>%
  
  #remove the duplicate entries with different senses/strands, but everyhting else is identical. 
  unique(.) #%>%

  ##For Checking on duplicate fusions
  # group_by(Sample) %>%
  # mutate(N_Fusion_Per_Sample=n()) %>% 
  # ungroup() %>% 
  # 
  # select(Sample,USI,N_Fusion_Per_Sample,
  #        everything()) 
  

```

```{r}
# dim(star.fmt) #3287   45
dim(star.fmt.filter) #
length(unique(star.fmt.filter$Sample)) #491 samples
```

```{r}
filter(star.fmt.filter,X.Fusion=="NoneDetected") #OK 5 Samples
# View(filter(star.fmt.filter,X.Fusion=="NonePassedFilter"))
```

```{r}
head(star.fmt.filter)
# write.csv(star.fmt.filter, "St.Jude_AML_ALL_STAR_Fusion_reformatted_FilteredForNBM_10.12.20.csv", row.names = FALSE)
```



#Define the Primary Fusions/Breakpoints per Sample 

```{r}
star.fmt.filter <- read.csv("St.Jude_AML_ALL_STAR_Fusion_reformatted_FilteredForNBM_10.12.20.csv")

dim(star.fmt.filter)
```


```{r warning=FALSE}
star.fmt.primary <- star.fmt.filter %>% 
  rowwise() %>%
  #Fusion category is to encompass both reciprocal fusions under 1 name. 
  mutate(Fusion.Category=fusionCategory(Left.Gene,Right.Gene)) %>% 
  mutate(JunctionRead = as.numeric(ifelse(grepl("None", JunctionRead),
                                          NA,JunctionRead)),
         SpanningRead = as.numeric(ifelse(grepl("None", SpanningRead),
                                          NA,SpanningRead))) %>%
  ungroup() %>%

  #Define the primary fusion per Sample per fusion. 
  group_by(Sample,Fusion.Category) %>%
  mutate(Primary.Fusion=getPrimaryFusion(USI=Sample, 
                                         junction_reads = JunctionRead, 
                                         spanning_reads = SpanningRead)) %>%
  arrange(Sample,Fusion.Category) %>%
  
  #Create a column for alterntate fusion breakpoints per Sample per fusion. 
  group_by(Sample,Fusion.Category,Primary.Fusion) %>%
  mutate(Alternate.Breakpoints=ifelse(Primary.Fusion=="alternate",
                                      collapseFusions(breakpoint = Breakpoints), NA)) %>%
  ungroup() %>%

  group_by(Sample,Fusion.Category) %>%
  mutate(Alternate.Breakpoints=fillNA(Alternate.Breakpoints, USI=Sample)) %>%
  mutate(Alternate.Breakpoints=ifelse(Breakpoints==Alternate.Breakpoints,
                                      "",Alternate.Breakpoints)) %>%
  ungroup() %>%
  
  #Filter out the alternate fusion rows, so alternates only provide breakpoint information. 
  filter(grepl("primary", Primary.Fusion) |
           grepl("NonePassedFilter|NoneDetected", X.Fusion)) %>%
  mutate_at(vars(Fusion.Category),
            ~ifelse(grepl("NonePassedFilter-NonePassedFilter", .),
                                                "NonePassedFilter", .)) %>%
  select(Sample,subject_name,
         Type,Fusion.Category,
         X.Fusion,Breakpoints,Alternate.Breakpoints,
         SpanningRead, JunctionRead, 
         Left.Gene, Left.Chr, Left.Breakpoint,
         Right.Gene,Right.Chr, Right.Breakpoint,
         Fusion.Detected.STAR,
         everything(),
         -Primary.Fusion)

  
dim(star.fmt.primary) # 2589   47
length(unique(star.fmt.primary$Sample)) #491
```

```{r}
# star.fmt.primary %>%
#   group_by(X.Fusion) %>%
#   summarize(N=n()) %>%
#   ungroup() %>%
#   arrange(desc(N)) %>%
#   slice(1:100) %>%
#   View()

# star.fmt.primary %>% 
#   filter(grepl("NoneDetected|NonePassed", Fusion.Detected.STAR)) %>%
#   View()
# table(star.fmt.primary$N)


# sum(grepl("CBFB-MYH11|MYH11-CBFB", star.fmt.primary$Fusion.Category))
```

```{r}
head(star.fmt.primary)
write.csv(star.fmt.primary, "St.Jude_AML_ALL_STAR_Fusion_reformatted_FilteredForNBM_PrimaryFusions_10.12.20.csv", row.names=FALSE)

# write.csv(star.fmt.primary, file.path(PROJHOME,"2020.09.15_RNAseq_Fusion_Breakpoints/References/St.Jude_AML_ALL_STAR_Fusion_reformatted_FilteredForNBM_PrimaryFusions_10.12.20.csv"), 
#           row.names=FALSE)

```

```{r}
table(star.fmt.primary$Fusion.Category) %>%
  data.frame() %>%
  arrange(desc(Freq)) %>%
  head(., n=20)
```

A little concerning that the top 20 most prevalent hits are NOT AML known fusions....


#Split into intra and interchromosomal Fusion

```{r}
inter <- star.fmt.primary  %>%
  filter(grepl("interchromosomal|NoneDetected|NonePassedFilter", Type)) %>%
  select(-Fusion.Category)

dim(inter) #905  46
# write.csv(inter, "St.Jude_AML_ALL_STAR_Fusion_reformatted_InterChromosomal_FilteredForNBM_PrimaryFusions_10.12.20.csv", row.names=FALSE)
```

```{r}
intra <- star.fmt.primary %>% 
  filter(grepl("intrachromosomal|NoneDetected|NonePassedFilter", Type)) %>%
  select(-Fusion.Category)

dim(intra) #1703   46
# write.csv(intra, "St.Jude_AML_ALL_STAR_Fusion_reformatted_IntraChromosomal_FilteredForNBM_PrimaryFusions_10.12.20.csv", row.names=FALSE)
```


#Split into Intra/Inter Chromosomal Fusions and One Row per Sample. 

```{r}
table(star.fmt.primary$Type)
```

```{r}
star.type <- star.fmt.primary %>%
  group_by(Type,Sample) %>%
  mutate(All.Fusions.Found=collapseRows(Fusion.Category,uniq = TRUE)) %>%
  ungroup()
  
head(star.type)
```


```{r}
FusionGenes <- star.type %>%
  select(Sample,All.Fusions.Found,Type) %>%
  unique(.) %>%
  spread(Type,All.Fusions.Found) %>%
  select(-NonePassedFilter,-NoneDetected)
 
View(FusionGenes)
dim(FusionGenes)
```


```{r}
dim(FusionGenes) #1107    4
# length(unique(star.type$USI))
# write.csv(FusionGenes, "St.Jude_AML_ALL_STAR_Fusion_FilteredForNBM_ByType_10.12.20.csv", row.names = F)
```








#Session Information

```{r}
sessionInfo()
```


















