---
title: 'Concatenate STAR Fusion for Relapse Samples'
author: "Jenny Smith"
date: "April 17, 2020"
output: html_document
---


```{r setup}
library(knitr)
# dir.create(file.path(PROJHOME,"2020.02.13_STAR-Fusion_Relapse"))
knitr::opts_knit$set(root.dir = file.path(PROJHOME,"2020.02.13_STAR-Fusion_Relapse"))

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center', fig.width = 10, fig.height = 10)

options(stringsAsFactors = FALSE)
```

```{r message = FALSE, warning=FALSE}
library(stringr)
library(magrittr)
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
getwd()
```

```{r}
source(file.path(SCRIPTS,"conversion_scripts/Merge_Cat_FixDupIDs_Function.r"))
```


#Define Functions to be used

```{r}
collapseRows <- function(col, uniq=FALSE){
  #designed for dplyr so that col is the vector of the column name
  if (uniq){col <- unique(col)}
  
  collapsed <- ifelse(all(is.na(col)), NA, paste(col, collapse = "; "))
  return(collapsed)
}


removeDups <- function(col.rowwise){
  
  uniqued <- str_split(col.rowwise, "; ") %>% 
  unlist() %>%
  gsub(" ","",.) %>%
  unique() %>%
  paste(.,collapse = "; ")
  
  return(uniqued)
  
}

```


#Read in the Manifest

```{r}
manifest <- read.csv(file.path(TARGET,"SequencingDataMatrix/TARGET_AML_RBD_Relapse_Manifest_v4.csv"), row.names = 1) %>% 
  select(Sample.ID, Colnames, USI,everything())

head(manifest[,1:5])
dim(manifest) #543  43

table(manifest$Group)
```


# Download the files from S3

```{bash eval=FALSE}
cd "/fh/scratch/delete90/meshinchi_s/jlsmith3/Relapse"

aws s3 cp  --recursive --exclude "*" --include "*annotated.coding_effect" --include "*html"  s3://fh-pi-meshinchi-s/SR/starfusion/relapsed_AML/ .

```


#Concatenate the Patient Level Files 

```{r}
path <- file.path(SCRATCH,"jlsmith3/Relapse")

files <- dir(path=path,recursive = TRUE,
             full.names = TRUE,
             pattern="*tsv.annotated.coding_effect")

length(files) #578
head(files)
```

```{r}
regex="^.+Relapse/(T.+[0-9]R).+_/Fusion.+"
fusionCaller <- catRbind(filenames=files, regex=regex, header=TRUE) 

fusionCaller <- fusionCaller %>% 
  mutate(USI=str_split_fixed(Patient, "-", n=5)[,3]) %>% 
  select(Patient,USI, everything()) 



dim(fusionCaller)#4,878   33
head(fusionCaller[,1:5])
# tail(fusionCaller)
length(unique(fusionCaller$Sample)) #538
# write.csv(fusionCaller, "TARGET_AML_0531_1031_Relapse_STAR_Fusion_GRCh37.csv", row.names=FALSE)
# getwd()
```

```{r}
length(unique(manifest$Sample.ID))
length(unique(fusionCaller$Patient)) #538

#These are missing due to the fact that STAR identified 0 fusions.
#just add these into the dataset with none detected category 
setdiff(manifest$Sample.ID,fusionCaller$Patient)
```

```{ bash eval=FALSE}
samples=("TARGET-20-PARLVL-09A-01R"
"TARGET-20-PARYJU-04A-01R"
"TARGET-20-PATLZB-04A-01R" 
"TARGET-20-PAVDBT-04A-01R"
"TARGET-20-PAWHJI-04A-01R")

for sample in $(echo ${samples[*]} ) ; do res=$(grep "PRE.$sample" starfusion.txt | tr -s " " |cut -f 3 -d " ");  aws s3 cp s3://fh-pi-meshinchi-s/SR/starfusion/relapsed_AML/${res}star-fusion.fusion_predictions.abridged.coding_effect.tsv - ; done

#all empty files
```



#Read in the STAR Fusion Data

```{r}
star <- read.csv("TARGET_AML_0531_1031_Relapse_STAR_Fusion_GRCh37.csv", 
                 stringsAsFactors = FALSE) %>% 
  
  #Add in a row for patients that had 0 fusions detected from the RNA-seq
  #These patients had empty dataframes produced by STAR-Fusion
  add_row(Patient=setdiff(manifest$Sample.ID,.$Patient)) %>% 
  select(-USI)

length(unique(star$Patient)) #543
dim(star) #4883   32
head(star)
```



# Reformat STAR fusion to Include Custom Columns 

```{r}
star.fmt <- star %>% 
    arrange(Patient) %>%
    select(Patient, X.Fusion=X.FusionName,
           Left.Breakpoint=LeftBreakpoint,
           Right.Breakpoint=RightBreakpoint,
           JunctionRead=JunctionReadCount,
           SpanningRead=SpanningFragCount,
          everything()) %>%
    #replace dashes with a dot. for consistency withother datasets.
    mutate_at(vars(Patient,X.Fusion,LeftGene,RightGene),
              ~gsub("-", "\\.", .)) %>% 
    mutate(Left.Gene=str_split_fixed(LeftGene,"\\^",n=2)[,1],
           Right.Gene=str_split_fixed(RightGene,"\\^",n=2)[,1],
           Left.Chr=str_split_fixed(Left.Breakpoint, ":", n=3)[,1],
           Right.Chr=str_split_fixed(Right.Breakpoint,":", n=3)[,1],
           
           Left.Strand=str_split_fixed(Left.Breakpoint, ":", n=3)[,3],
           Right.Strand=str_split_fixed(Right.Breakpoint,":", n=3)[,3],
           
           Breakpoints = paste(gsub("chr|:[\\+-]","",Left.Breakpoint),
                               gsub("chr|:[\\+-]","",Right.Breakpoint),
                               sep="|"),
           Type = ifelse(Left.Chr == Right.Chr,
                         "intrachromosomal","interchromosomal")
           # size = ifelse(Type=="intrachromosomal", 
           #               abs(Left.Breakpoint - Right.Breakpoint), NA)
           ) %>%
    mutate(X.Fusion = gsub("\\.{2}","-",X.Fusion)) %>%
    mutate(Fusion.Detected.STAR=ifelse(is.na(X.Fusion),
                                       "NoneDetected","Detected")) %>%
    mutate_at(vars(X.Fusion:Type),
              ~ifelse(Fusion.Detected.STAR=="NoneDetected",
                      "NoneDetected",.)) %>%
    mutate(USI=str_split_fixed(Patient, "\\.", n=5)[,3]) %>%
    select(Patient,USI,everything())
  

dim(star.fmt)
length(unique(star.fmt$Patient)) 
```

```{r}
head(star.fmt)
# write.csv(star.fmt, "TARGET_AML_0531_1031_Relapse_STAR_Fusion_reformatted_4.22.2020.csv", row.names = FALSE)
```



#Filter out Fusions found in NBMs 

```{r}
filterFusions <- function(X.Fusion, Breakpoints,df2.Fusion,df2.Breakpoints){

  fus <- paste(X.Fusion,Breakpoints,sep=" ")
  # print(fus)
  fus2 <- paste(df2.Fusion, df2.Breakpoints, sep=" ")
  
  if (fus %in% fus2){
      res <- 1 #these fusions will be filtered out. 
  }else{
    res <- 0 #these one will not. 
  }
  
  return(res)
}
```

```{r}
NBM.1031 <- read.csv(file.path(PROJHOME,"2018.02.07_STAR-Fusion_C.Nguyen","Reformatted/TARGET_AML_1031_STAR_Fusion_reformatted_7.02.2018.csv"), 
                     row.names = 1) %>%
  filter(grepl("BM[0-9]|RO[0-9]", USI)) %>% 
  mutate_all(~as.character(.))


head(NBM.1031)
dim(NBM.1031) #5707   17
```

```{r}
# setdiff(colnames(NBM.1031), colnames(star.fmt)) #size only
setdiff(colnames(star.fmt),colnames(NBM.1031))
```

```{r}
NBM.0531 <- read.csv(file.path(PROJHOME,"2018.02.07_STAR-Fusion_C.Nguyen","Reformatted/TARGET_AML_0531_STAR_Fusion_reformatted_4.09.2019.csv")) %>%
  filter(grepl("BM[0-9]|RO[0-9]", USI)) %>% 
  mutate_all(~as.character(.))


head(NBM.0531)
dim(NBM.0531) #1757   19
```

```{r}
#Using both NBM sets to account for the total fusions found in the NBM dataset
NBM <- star.fmt %>%
  filter(grepl("BM[0-9]|R[O0][0-9]", USI)) %>%
  mutate_all(~as.character(.)) %>%
  bind_rows(., NBM.1031,NBM.0531) %>% 
  select(-one_of(setdiff(colnames(star.fmt),colnames(NBM.1031))))

# head(NBM)
dim(NBM)  # 7583   45 (119 in CD34+ NBM alone)
length(unique(NBM$Patient)) #100
# tail(NBM)

# write.csv(NBM,"~/TARGET_Fusions_in_NormalMarrrows_STAR-fusion_4.27.20.csv",
#           row.names = FALSE)
```

```{r}
star.fmt.filter <- star.fmt %>% 
  select(Patient,USI,everything()) %>%
  
  rowwise() %>%
  mutate(FailedFilterFusions=filterFusions(X.Fusion, 
                                           Breakpoints,
                                           NBM$X.Fusion,
                                           NBM$Breakpoints)) %>%
  ungroup() %>%
    
  group_by(Patient) %>%
  mutate(AllFail=ifelse(sum(FailedFilterFusions) == n(),"Yes","No")) %>%
  mutate(Fusion.Detected.STAR=ifelse(AllFail == "Yes" ,
                                     "NonePassedFilter",
                                     Fusion.Detected.STAR)) %>%
  ungroup()  %>%
  
  filter(AllFail == "Yes" | FailedFilterFusions == 0) %>%

  rowwise() %>%
  #change all to characters so that numeric cols can be filled below with a charcter string.
  mutate_all(funs(as.character(.))) %>% 
  mutate_at(vars(X.Fusion:Fusion.Detected.STAR), 
            ~ifelse(Fusion.Detected.STAR =="NonePassedFilter",
                    "NonePassedFilter",  .)) %>%
  ungroup() %>%
  
  select(Patient,USI,
         everything(),-AllFail,-FailedFilterFusions) %>%
          # -Left.Strand,-Right.Strand
  
  #remove the duplicate entries with different senses/strands, but everyhting else is identical. 
  unique(.) #%>%

  ##For Checking on duplicate fusions
  # group_by(Patient) %>%
  # mutate(N_Fusion_Per_Patient=n()) %>% 
  # ungroup() %>% 
  # 
  # select(Patient,USI,N_Fusion_Per_Patient,
  #        everything()) 
  

```

```{r}
# dim(star.fmt) #4883   42
dim(star.fmt.filter) #2003    40
length(unique(star.fmt.filter$Patient)) #543 samples
```

```{r}
# filter(star.fmt.filter,X.Fusion=="NoneDetected") #OK 5 patients
# View(filter(star.fmt.filter,X.Fusion=="NonePassedFilter"))
```

```{r}
# arrange(star.fmt.filter, desc(Patient),
#         desc(N_Fusion_Per_Patient),
#         X.Fusion) %>% 
#   slice(1:50) %>%
#   View()
```

```{r}
head(star.fmt.filter)
# write.csv(star.fmt.filter, "TARGET_AML_0531_1031_Relapse_STAR_Fusion_reformatted_FilteredForNBM_4.30.2020.csv", row.names = FALSE)
```



#Define the Primary Fusions/Breakpoints per patient 

```{r}
fusionCategory <- function(geneA,geneB){
  fus <- c(geneA,geneB)
  fus <- paste(fus[order(fus)], collapse = "-") #order alphabetically, so that the category includes  reciprocal fusions
}
```

```{r}
collapseFusions <- function(breakpoint){
  paste(unique(breakpoint), collapse = "; ")
}
```

```{r}
fillNA <- function(x,USI){
  
  USI <- unique(USI)
  x <- x %>% .[!is.na(.)] %>% unique()
  
  if(length(x) < 1){
    x <- ""
  }
  
  # print(x)
  return(x)
}
```

```{r}
getPrimaryFusion <- function(USI,junction_reads,spanning_reads){


  resolveTies <- function(df){
        df <- df %>%
            group_by(primary) %>%
            #if still a tie in breakpoint pairs counts, choose the first.
            mutate(rank=rank(junction_reads, ties.method = "first")) %>%  
            #convert all alternate fusion calls to zero rank (eg not going to be used in tie-breaking)
            mutate(rank=ifelse(primary=="alternate", 0, rank)) %>%
            rowwise() %>%
            mutate(reason=sum(reason,rank)) %>%
            ungroup() %>% 
            #the fusion with highest # of breakpoint reads will have largest interger rank. 
            mutate(primary=ifelse(reason == max(reason),
                                  "primary", "alternate"))
        
        return(df)
    }
  
  #First check that the reads are not NA values (eg. NBM samples)
  if (all(is.na(junction_reads))){
    res <- NA
    return(res)
  }
  
  
  #Create boolean vectors for each type of evidence supporting the "primary" fusion. 
  USI <- unique(USI)
  brk.reads <-  junction_reads == max(junction_reads)
  spanning.reads <- spanning_reads == max(spanning_reads)

  #Create a temporary data frame. 
  temp <- data.frame("brk.reads.bool"=brk.reads,
                     "spanning.reads.bool"=spanning.reads, 
                     stringsAsFactors = FALSE) %>%
    #add in the subsetted per patient read counts. For visual checking on ties, etc. 
    bind_cols(.,list("junction_reads"=junction_reads,
                     "spanning_reads"=spanning_reads))
  
  #will select the one with the highest reads
    temp <- temp %>%
        rowwise() %>%
        mutate(reason=sum(brk.reads.bool,spanning.reads.bool)) %>%
        ungroup() %>%
        mutate(primary=ifelse(reason == max(reason), "primary", "alternate"))

    # break ties if there are multiple fusions with same #s of supporting reads.
    # for those with highest evidence ("reason"), choose the one with the higher breakpoint reads
    if (sum(temp$primary == "primary") > 1){
        temp <- resolveTies(temp)
        
        # last check for ties, will break dplyr by returning a dataframe, not a vector. 
        if (sum(temp$primary == "primary") > 1 ){
          return(temp)
        }
    }

  res <- temp$primary
  
  return(res)
}
```


```{r warning=FALSE}
star.fmt.primary <- star.fmt.filter %>% 
  rowwise() %>%
  #Fusion category is to encompass both reciprocal fusions under 1 name. 
  mutate(Fusion.Category=fusionCategory(Left.Gene,Right.Gene)) %>% 
  mutate(JunctionRead = as.numeric(ifelse(grepl("None", JunctionRead),
                                          NA,JunctionRead)),
         SpanningRead = as.numeric(ifelse(grepl("None", SpanningRead),
                                          NA,SpanningRead)), 
         Group=case_when(
           grepl("PAXLWH", Patient) ~ "AML", #cell cultured AML sample
           grepl("BM[0-9].+34POS|R[O0][0-9].+34POS", Patient)~ "CD34_NBM",
           grepl("BM[0-9]{2,}|R[O0][0-9]", Patient)~ "NBM",
           grepl("MPN[0-9]", Patient) ~ "MPN",
           grepl("sort",Patient, ignore.case = T) ~ "FlowSorted",
           grepl("CSGH|K562|NOMO|ME1|MO7E", USI) ~ "CellLine",
           TRUE ~ "AML"),
         Time_point=case_when(
           grepl("09A|03A", Patient) ~ "diagnostic",
           grepl("04A|40A", Patient) ~ "relapse",
           TRUE ~ Group)) %>%
  ungroup() %>%

  #Define the primary fusion per patient per fusion. 
  group_by(Patient,Fusion.Category) %>%
  mutate(Primary.Fusion=getPrimaryFusion(USI=Patient, 
                                         junction_reads = JunctionRead, 
                                         spanning_reads = SpanningRead)) %>%
  arrange(Patient,Fusion.Category) %>%
  
  #Create a column for alterntate fusion breakpoints per patient per fusion. 
  group_by(Patient,Fusion.Category,Primary.Fusion) %>%
  mutate(Alternate.Breakpoints=ifelse(Primary.Fusion=="alternate",
                                      collapseFusions(breakpoint = Breakpoints), NA)) %>%
  ungroup() %>%

  group_by(Patient,Fusion.Category) %>%
  mutate(Alternate.Breakpoints=fillNA(Alternate.Breakpoints, USI=Patient)) %>%
  mutate(Alternate.Breakpoints=ifelse(Breakpoints==Alternate.Breakpoints,
                                      "",Alternate.Breakpoints)) %>%
  ungroup() %>%
  
  #Filter out the alternate fusion rows, so alternates only provide breakpoint information. 
  filter(grepl("primary", Primary.Fusion) |
           grepl("NonePassedFilter|NoneDetected", X.Fusion)) %>%
  mutate_at(vars(Fusion.Category),
            ~ifelse(grepl("NonePassedFilter-NonePassedFilter", .),
                                                "NonePassedFilter", .)) %>%
  select(Patient,USI,Group,Time_point,
         Type,Fusion.Category,
         X.Fusion,Breakpoints,Alternate.Breakpoints,
         SpanningRead, JunctionRead, 
         Left.Gene, Left.Chr, Left.Breakpoint,
         Right.Gene,Right.Chr, Right.Breakpoint,
         Fusion.Detected.STAR,
         everything(),
         -Primary.Fusion)

  
dim(star.fmt.primary) #1349   45
length(unique(star.fmt.primary$Patient)) #543
table(star.fmt.primary$Group,useNA='ifany')
table(star.fmt.primary$Time_point,useNA='ifany')
```

```{r}
# star.fmt.primary %>% 
#   select(Patient,Group, X.Fusion) %>% 
#   group_by(Patient,X.Fusion) %>% 
#   summarize(N=n()) %>% 
#   ungroup() %>% 
#   arrange(desc(N)) %>% 
#   slice(1:100) %>%
#   View()

# star.fmt.primary %>% 
#   filter(grepl("NoneDetected|NonePassed", Fusion.Detected.STAR)) %>%
#   View()
# table(star.fmt.primary$N)
```

```{r}
head(star.fmt.primary)
# write.csv(star.fmt.primary, "TARGET_AML_0531_1031_Relapse_STAR_Fusion_reformatted_FilteredForNBM_PrimaryFusions_5.01.2020.csv", row.names=FALSE)
```

```{r}
table(star.fmt.primary$Fusion.Category) %>%
  data.frame() %>%
  arrange(desc(Freq)) %>%
  head(., n=20)
```

A little concerning that the top 20 most prevalent hits are NOT AML known fusions....


#Split into intra and interchromosomal Fusion

```{r}
inter <- star.fmt.primary  %>%
  filter(grepl("interchromosomal|NoneDetected|NonePassedFilter", Type)) %>%
  select(-Fusion.Category)

dim(inter) #624  45
# write.csv(inter, "TARGET_AML_0531_1031_Relapse_STAR_Fusion_reformatted_InterChromosomal_FilteredForNBM_PrimaryFusions_5.01.2020.csv", row.names=FALSE)
```

```{r}
intra <- star.fmt.primary %>% 
  filter(grepl("intrachromosomal|NoneDetected|NonePassedFilter", Type)) %>%
  select(-Fusion.Category)

dim(intra) #879  45
# write.csv(intra, "TARGET_AML_0531_1031_Relapse_STAR_Fusion_reformatted_IntraChromosomal_FilteredForNBM_PrimaryFusions_5.01.2020.csv", row.names=FALSE)
```


#Split into Intra/Inter Chromosomal Fusions and One Row per Patient. 

```{r}
table(star.fmt.primary$Type)
```

```{r}
star.type <- star.fmt.primary %>%
  group_by(Type,Patient) %>%
  mutate(All.Fusions.Found=collapseRows(Fusion.Category,uniq = TRUE)) %>%
  ungroup()
  
# head(star.type)
```


```{r}
FusionGenes <- star.type %>%
  select(Patient,All.Fusions.Found,Type) %>%
  unique(.) %>%
  spread(Type,All.Fusions.Found) %>%
  
  mutate(Group=case_when(
           grepl("PAXLWH", Patient) ~ "AML", #cell cultured AML sample
           grepl("BM[0-9].+34POS|R[O0][0-9].+34POS", Patient)~ "CD34_NBM",
           grepl("BM[0-9]{2,}|R[O0][0-9]", Patient)~ "NBM",
           grepl("MPN[0-9]", Patient) ~ "MPN",
           grepl("sort",Patient, ignore.case = T) ~ "FlowSorted",
           grepl("CSGH|K562|NOMO|ME1|MO7E", Patient) ~ "CellLine",
           TRUE ~ "AML"),
         Time_point=case_when(
           grepl("09A|03A", Patient) ~ "diagnostic",
           grepl("04A|40A", Patient) ~ "relapse",
           TRUE ~ Group)) %>% 
  mutate_at(vars(interchromosomal:intrachromosomal), 
            ~case_when( 
            is.na(.) & Group != "CD34_NBM" ~ "NoneDetected", 
            is.na(.) & Group == "CD34_NBM" ~ "NonePassedFilter", 
            TRUE ~ .)) %>%
  select(-NonePassedFilter)
 
# View(FusionGenes)
```


```{r}
dim(FusionGenes) #1107    4
# length(unique(star.type$USI))
# write.csv(FusionGenes, "TARGET_AML_0531_STAR_Fusion_FilteredForNBM_ByType_7.02.2018.csv")
```


#Tally the Fusions 


```{r}
createMultiCols <- function(col,split=FALSE,suffix){
  #col is the column with many factors (eg all fusions). character vector lenght 1.
  #suffix is a character string for the suffix on all column names
  #desinged for dplyr bind_cols() function
  
  #example
  # df <- df %>%
  #   bind_cols(.,createMultiCols(.$Fusion, suffix=".sensefusion")) 
  
  if(split){
    groups <- unique(gsub(" ","",unlist(str_split(col, "; "))))
    groups <- groups[groups != ""]
  }else{
    groups <- unique(col)
  }
  list <- lapply(groups, function(x) ifelse(grepl(paste0(x, ";"), col) | grepl(paste0(x, "$"), col) , "Yes",
                                            ifelse(grepl("^$|Unknown",col) | is.na(col), "Unknown","No"))) 
  list.names <- gsub("-", "_",  groups)
  names(list) <- paste0(list.names,suffix)
  
  return(list)
}
```



```{r}
createCols_byType <- function(df,col){
  
  # print(head(unlist(df[,col])))
  
  cols <- createMultiCols(col=unlist(df[,col]),split = TRUE,suffix = c(""))  %>%
    bind_cols() %>%
    mutate(USI=unlist(df[,"USI"])) %>%
    select(USI, everything(), -NonePassedFilter, -NoneDetected)
  
  return(cols)
}
```


```{r}
createFreqTable <- function(createMultiCols.df,type){
  df <- t(sapply(createMultiCols.df[,-1], table)) %>%
          as.data.frame() %>% 
          rownames_to_column("Fusion_Category") %>%
          mutate(Percent=(Yes/(Yes+No))*100) %>%
          arrange(desc(Percent))

  if (type == "Exon"){
    df <- df %>%
            separate(Fusion,c("X5.Gene", "X3.Gene"), sep="exon[0-9]{1,2}", remove = FALSE) %>%
            mutate(X5.Exon=gsub("^.+(exon[0-9]{1,2})_.+", "\\1", Fusion),
                 X3.Exon=gsub("^.+_.+(exon[0-9]{1,2})", "\\1", Fusion),
                 X3.Gene=gsub("_", "", X3.Gene)) %>%
            select(Fusion,contains("Gene"),contains("Exon"),everything())

  }else if(type == "Gene"){
    df <- df %>%
            separate(Fusion_Category,c("GeneA", "GeneB"), sep="\\_", remove = FALSE) %>%
            select(Fusion_Category,contains("Gene"),everything())
  }
  
  return(df)
}
```


```{r}
inter.Genes <- createCols_byType(df=FusionGenes, col="interchromosomal") 
dim(inter.Genes) #1112 by 17,126
```


```{r warning=FALSE}
Inter.G.Tab <- createFreqTable(inter.Genes, type="Gene")
head(Inter.G.Tab)
# write.csv(Inter.G.Tab, "TARGET_AML_0531_STAR_Fusion_reformatted_InterChromosomal_FilteredForNBM_GeneLevel_Tables_07.02.2018.csv", row.names = FALSE)
```


```{r}
intra.Genes <- createCols_byType(df=FusionGenes, col="intrachromosomal")

dim(intra.Genes) #1112 by 10,837
```


```{r warning=FALSE}
Intra.G.Tab <- createFreqTable(intra.Genes, type="Gene")
head(Intra.G.Tab)

# write.csv(Intra.G.Tab, "TARGET_AML_0531_STAR_Fusion_reformatted_IntraChromosomal_FilteredForNBM_GeneLevel_Tables.csv", row.names = FALSE)
```



#Reformat Star for AGFusion 

Will need the in-frame information and exon information. 

```{r}
star.fmt.primary <- read.csv("Reformatted/TARGET_AML_0531_STAR_Fusion_reformatted_FilteredForNBM_PrimaryFusions_7.02.2018.csv", row.names = 1)
dim(star.fmt.primary)
head(star.fmt.primary[,1:5])
```

STAR fusion format does not work because the AGFusion expects to have an ENSG identifier, which I do not have. Though these may be 

```{r}
# ForAnno <- star %>%
#   # slice(1:1000) %>%
#   select(FusionName=X.Fusion, JunctionReadCount=JunctionRead, SpanningFragCount=SpanningRead,
#          LeftGene=Left.Gene, Left.Chr, Left.Breakpoint, Left.Strand,
#          RightGene=Right.Gene, Right.Chr, Right.Breakpoint, Right.Strand) %>%
#   mutate(SpliceType=rep("None", nrow(.)), 
#          FusionName=gsub("_","--",FusionName), 
#          LargeAnchorSupport=rep("None",nrow(.))) %>%
#   unite(LeftBreakpoint, Left.Chr:Left.Strand, sep=":") %>%
#   unite(RightBreakpoint, Right.Chr:Right.Strand, sep=":") %>%
#   select(FusionName,JunctionReadCount, SpanningFragCount, SpliceType, LeftGene, LeftBreakpoint,
#          RightGene, RightBreakpoint,LargeAnchorSupport)
# 
# ForAnno  
```


So I reformatted the data to be like the Tophat input. 
Cu sent over the GTF file, and the STAR fusion used Ensembl V74 annotations. 

```{r}
ForAnno <- star %>% 
  # slice(1:100) %>%
 select(Patient, Left.Gene, Left.Chr, Left.Breakpoint, 
         Right.Gene, Right.Chr, Right.Breakpoint, 
        JunctionRead, SpanningRead) %>% 
  mutate(Col10=rep(0,nrow(.)), 
         Col11=rep(0,nrow(.))) 
  #remove the version #s from the ACXXXX genes. 
  # mutate_at(vars(contains("Gene")), funs(ifelse(grepl("^AC[0-9]+", .), gsub("\\.[0-9]", "",.), .))) 
  # filter(grepl("SNX29P2", Left.Gene) | grepl("SNX29P2", Right.Gene))
```

```{r}
# write.table(ForAnno, "TARGET_AML_0531_Starfusion_inTopHapFmt_ForAGFusion.txt", 
#             row.names = FALSE, sep = "\t", quote = FALSE, col.names =FALSE)
```







#Session Information

```{r}
sessionInfo()
```


















