---
title: "Concatenate RNA-seq"
author: "Jenny Smith"
date: "June 6, 2019"
output: html_document
---
  

#Set-up
  
```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center', fig.height = 5, fig.width = 5)
knitr::opts_knit$set(root.dir = file.path(PROJHOME,'/2019.06.06_Concatenate_Relapse_RNAseq/'))
options(stringsAsFactors = FALSE)
```

```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(tidyr)
library(tibble)
library(magrittr)
library(purrr)
library(readr)
```

```{r}
source(file.path(SCRIPTS,"conversion_scripts/Merge_Cat_FixDupIDs_Function.r"))
source(file.path(SCRIPTS,"RNAseq_Analysis/DifferentialExpn_PathwayAnalysis/R/rmDupGenes_Function.r"))
```


#Define Functions to be used. 

```{r}
#Function for the TPM conversion. 
# Based on https://groups.google.com/forum/#!topic/rsem-users/W9RQrZIOzA4
#Useage: sapply(cated$RPKM, RPKM_to_TPM)

RPKM_to_TPM <- function(RPKM){
  conversionFactor <- sum(RPKM) / 1E6
  TPM <- RPKM / conversionFactor
  return(TPM)
}
```


```{r}
#https://stackoverflow.com/questions/32040742/dplyrselect-including-all-other-columns-at-end-of-new-data-frame-or-beginni

addCols <- function(df,symbol,feature_number=NULL){
  library(dplyr)
  
  df <- df %>%
    as.data.frame() %>%
    mutate(geneSymbol=symbol) %>%
    select(geneSymbol, everything()) 
  
  if(!is.null(feature_number)){
    df <- df %>% 
      add_column(feature_number=feature_number,.after = "geneSymbol")
  }
  
  return(df)
}
```



#Read in the Manifests and ID maps

```{r}
ID.map1 <- read.csv("target_barcode_mapping_4.8.19.csv") %>% 
  mutate(Batch="1") %>% 
  set_rownames(.$Barcode)

dim(ID.map1)
head(ID.map1)
```

```{r}
ID.map2 <- read.csv("target_barcode_mapping_5.31.19.csv") %>% 
  mutate(Batch="2") %>% 
  set_rownames(.$Barcode)

dim(ID.map2)
head(ID.map2)
```

```{r}
ID.map3 <- read.csv("target_barcode_mapping_6.14.19.csv") %>% 
  mutate(Batch="3") %>% 
  set_rownames(.$Barcode)

dim(ID.map3)
head(ID.map3)
```

```{r}
ID.map4 <- read.csv("target_barcode_mapping_9.17.19.csv") %>%
    select(GSC.library=Library, Barcode=External.identifier, everything()) %>%
    mutate(Batch="4") %>% 
    set_rownames(.$Barcode)

dim(ID.map4)
head(ID.map4)
```


```{r}
manifest <- read.csv("GSC_1669_complete_manifest.csv", fileEncoding="latin1") %>% 
  left_join(., ID.map1, by=c("Sample.ID"="Barcode")) %>% 
  left_join(.,ID.map2, by=c("Sample.ID"="Barcode")) %>% 
  left_join(.,ID.map3, by=c("Sample.ID"="Barcode")) %>%
  left_join(.,ID.map4, by=c("Sample.ID"="Barcode")) %>%
  mutate_at(vars(matches("GSC.library|Batch")), ~ifelse(is.na(.), "", .)) %>% 
  unite(GSC.library,contains("GSC.library"), sep = "") %>%
  unite(Batch,contains("Batch"), sep = "") %>% 
  mutate(USI=str_split_fixed(Anonymous.Patient.ID, "-", n=3)[,3])

dim(manifest) #551  33
head(manifest)
```

```{r}
sum(manifest$GSC.library=="")
table(manifest$Batch, useNA = 'always')
# write.csv(manifest, "/fh/fast/meshinchi_s/workingDir/TARGET/AML_TARGET/SequencingDataMatrix/TARGET_AML_RBD_Relapse_Manifest_v3.csv")
```



#Identify Files to Be Concatenated

We have intron and exon level counts. However, I would want only the concatenated level files to be on the fast drive now. All patient level can be archived directly. 

```{r}
filepath <-  file.path(SCRATCH,"2019.04_BCCA_mRNAseq_Relapse_Data_Downloads")

gene <- dir(path = filepath, 
                pattern = "transcript.normalized$",
                recursive = TRUE,
                full.names = TRUE)

head(gene)
length(gene) #543
```

```{r}
intron <- dir(path = filepath, pattern = "intron.ratios.normalized$",
                recursive = TRUE,full.names = TRUE)

head(intron)
length(intron)  #543
```

```{r}
exon <- dir(path = filepath, pattern = "exon.normalized$",
                recursive = TRUE,full.names = TRUE)

head(exon)
length(exon) #543
```



#Columns Description

the .transcript.normalized file contains:
-*column 1: gene id*
-column 2: transcript id or "merged_<gene_id>" if the analysis was performed in
collapse mode
-colomn 3: chromosome
-column 4: start of the transcript (relative to the positive strand, so
start<end)
-column 5: end of the transcript (relative to the positive strand, so start<end)
-column 6: strand of the annotation ("+" or "-")
-column 7: sum of the length of all exons in this transcript or collapsed gene
-column 8: sum of the length of all introns in this transcript or collapsed gene
-*column 9*: number of fractional reads inside this merged_gene or transcript's
exons (sum of the fraction of each read inside all exons)
-column 10: total coverage across all the exons in this merged_gene or transcript
(sum of the coverage depth at each base in all exons)
-column 11: number of fractional reads inside this merged_gene or transcript's
introns (sum of the fraction of each read inside all introns)
-column 12: total coverage across all the introns in this merged_gene or
transcript (sum of the coverage depth at each base in all introns)
-column 13: average coverage over all exons -- sum of the coverage depth at each
base in all exons divided by the sum of the exon lengths)
-column 14: average coverage over all introns and exons -- sum of the coverage
depth at each base between the merged_gene or transcript's start and end
divided by the number of bases between the gene's start and end
-*column 15*: normalized coverage (RPKM) -- (number of fractional reads in all exons
in this merged gene or transcript x 1000000000)/(NORM_TOTAL x sum of the
length of all exons in this merged gene or transcript)
-*column 16*: gene symbol (or gene id if symbol is unavailable)
-column 17: biotype (or "-" if biotype unavailable)
-column 18: gene description (or "-" if description is unavailable)


_The .exon.normalized file contains: For BCCA_
-*column 1*: gene id
-column 2: transcript id or "merged_<gene_id>" if the analysis was performed in
collapse mode
-colomn 3: strand of the annotation ("+" or "-")
-*column 4*: exon number (relative to the gene's strand)
-column 5: "exon" (annotation type)
-column 6: chromosome
-column 7: start of the exon (relative to the positive strand, so start<end)
-column 8: end of the exon (relative to the positive strand, so start<end)
-*column 9*: number of fractional reads inside this exon's boundaries (sum of the
fraction of each read inside this region)
-column 10: total coverage across this exon (sum of the coverage depth at each
base in this exon)
-column 11: average coverage inside this exon-- total coverage across this exon
divided by the exon's length
-*column 12*: normalized average coverage inside this exon (RPKM) -- (number of this
exon's fractional reads x 1000000000) / (NORM_TOTAL x exon length)


_The .intron.ratios.normalized file contains:_
-*column 1*: gene id
-column 2: transcript id or "merged_<gene_id>" if the analysis was performed in
collapse mode
-colomn 3: strand of the annotation ("+" or "-")
-*column 4*: intron number (relative to the gene's strand)
-column 5: "intron" (annotation type)
-column 6: chromosome
-column 7: start of the intron (relative to the positive strand, so start<end)
-column 8: end of the intron (relative to the positive strand, so start<end)
-*column 9*: number of fractional reads inside this intron's boundaries (sum of the
fraction of each read inside this region)
-column 10: total coverage across this intron (sum of the coverage depth at each
base in this intron)
-column 11: collapsed gene or transcript length
-column 12: average coverage of the collapsed gene or transcript
-column 13: intron length
-column 14: average coverage inside this intron-- total coverage across this
intron divided by the intron's length
-column 15: intron-to-gene coverage ratio: average intron coverage divided by the
average collapsed gene or transcript coverage
-column 16: duplicate of column 14
-*column 17*: normalized average coverage inside this intron (RPKM) -- (number of
this intron's fractional reads x 1000000000) / (NORM_TOTAL x intron length)


#Begin Concatenation 

```{r}
results <- list("gene"=gene,"exon"=exon,"intron"=intron)
```

```{r}
#Pattern to select the Target Barcode/GSC barcode
pattern <- "^.+\\/(B[0-9]{5})_.+"
```

```{r}
#Select the column indices 
selected <- list("gene"=c(1,9,15,16),
                 "exon"=c(1,4,9,12),
                 "intron"=c(1,4,9,17)) 
```

```{r}
cated <- map2(results, selected, catExpnData, regex=pattern)
names(cated) <- names(results)

names(cated$gene) <- c("gene_id","counts","RPKM","geneSymbol")
names(cated$exon) <- c("gene_id","exon_number","counts","RPKM")
names(cated$intron) <- c("gene_id","intron_number","counts","RPKM")
```

```{r}
summary(cated)
```


#Add in TPMs 

```{r}
cated$gene[["TPM"]] <- apply(cated$gene$RPKM,2, RPKM_to_TPM)

table(apply(cated$gene[["TPM"]], 2, sum)) #all sum to 1 million 
```

```{r}
cated$exon[["TPM"]] <- apply(cated$exon$RPKM,2, RPKM_to_TPM)

table(apply(cated$exon[["TPM"]], 2, sum)) #all sum to 1 million 
```


```{r}
cated$intron[["TPM"]] <- apply(cated$intron$RPKM,2, RPKM_to_TPM)

table(apply(cated$intron[["TPM"]], 2, sum)) #all sum to 1 million 
```

```{r}
map(cated, ~map(., ~head(.[,1:5])))
```

```{r}
map(cated, ~map(., dim))
```


#Check that the Gene Ids are in the Same Order

```{r}
table(apply(cated$gene$geneSymbol,2,
            function(x) identical(x=x, y=cated$gene$geneSymbol[,1])))
```

```{r}
table(apply(cated$exon$gene_id,2, 
            function(x) identical(x=x, y=cated$exon$gene_id[,1])))
```

```{r}
table(apply(cated$intron$gene_id,2, 
            function(x) identical(x=x, y=cated$intron$gene_id[,1])))
```


#Add columns for the Gene Symbol and Ensembl Symbol 

```{r}
#Rename based on the provided Manifest 
cols <- data.frame(GSC.library=colnames(cated$gene$counts)) %>% 
  left_join(., select(manifest, Sample.ID, GSC.library), 
            by="GSC.library") #ensure correct order 

head(cols)
dim(cols)
```

```{r}
gene.final <- lapply(cated$gene[c(2,3,5)], 
                     function(x) addCols(df=x, symbol = cated$gene$geneSymbol[,1]) %>% 
                       set_colnames(., value = c("geneSymbol", gsub("-",".",cols$Sample.ID))))


lapply(gene.final,function(x) head(x[,1:5]))
# map(gene.final, ~sum(is.na(colnames(.)))) #no NAs
# map(gene.final, ~sum(duplicated(colnames(.)))) #no dups
```

```{r}
exon.final <- lapply(cated$exon[c(3:5)], function(x) addCols(df=x, symbol = cated$exon$gene_id[,1], 
                                                             feature_number = cated$exon$exon_number[,1]) %>% 
                       set_colnames(., value = c("gene_id", "exon_number", gsub("-",".",cols$Sample.ID))))

lapply(exon.final,function(x) head(x[,1:5]))
```

```{r}
intron.final <- lapply(cated$intron[c(3:5)], function(x) addCols(df=x, symbol = cated$intron$gene_id[,1], 
                                                          feature_number = cated$intron$intron_number[,1]) %>% 
                         set_colnames(., value = c("gene_id","intron_number", gsub("-",".",cols$Sample.ID))))


lapply(intron.final,function(x) head(x[,1:5]))
```

#Save the output

```{r}
for (final in ls(pattern="final")){
  for (df in names(get(final))){
      if(df == "counts"){
        n <- "FractionalCounts"
      }else{
        n <- df
      }
    
      type <- gsub(".final","Level",final)
      filename <- paste("TARGET_0531_1031_Relapse_Ribodepleted_RNAseq",type,n,sep="_") %>% 
        paste0(.,".csv")
      
      print(filename)
      write.csv(get(final)[[df]], file = filename, row.names = FALSE)
  }
}

```


#Remove Duplicate Genes 

```{r}
cts <- read_csv("ExpnData/geneLevel/TARGET_0531_1031_Relapse_Ribodepleted_RNAseq_geneLevel_FractionalCounts.csv")
cts <- as.data.frame(cts)

dim(cts)
head(cts[,1:5])
```

```{r}
TPM <- read_csv("ExpnData/geneLevel/TARGET_0531_1031_Relapse_Ribodepleted_RNAseq_geneLevel_TPM.csv")
TPM <- as.data.frame(TPM)

dim(TPM)
head(TPM[,1:5])
```


```{r}
rmDup_cts <- rmDupGenes(expnData = cts, 
                        geneCol = "geneSymbol")

dim(rmDup_cts$remDups) #51573   543
head(rmDup_cts$remDups[,1:5])


write.csv(rmDup_cts$remDups,"ExpnData/geneLevel/TARGET_0531_1031_Relapse_Ribodepleted_RNAseq_geneLevel_dupGenesRemoved_FractionalCounts.csv")
```

```{r}
rmDup_TPM <- rmDupGenes(expnData = TPM, 
                        geneCol="geneSymbol")

dim(rmDup_TPM$remDups) #51573   543
head(rmDup_TPM$remDups[,1:5])

# write.csv(rmDup_TPM$remDups, "ExpnData/geneLevel/TARGET_0531_1031_Relapse_Ribodepleted_RNAseq_geneLevel_dupGenesRemoved_TPM.csv")
```



NOTE:
  
  If markdown cannot be completed with pandoc (memory issues), use the following

> library("rmarkdown")
> render("~/scripts/RNAseq_Analysis/DifferentialExpn_PathwayAnalysis/TARGET_AML_1031_Cat_RNAseq.Rmd", "html_document")


#Session Info

```{r}
sessionInfo()
```

